\normalsize{A general approach and tool to evaluate the effectiveness of no-take marine reserves}

### 2017-08-18 Manuscript.Rmd
Updated results

# Results and Discussion

The results shown here intend to highlight the relevance and utility of the framework and tool, which automate the analysis and make it replicable. First, we present the scorecard generated by MAREA to provide an overview of the state of the reserve. Then, we provide (mdoified for style) a set of figures generated by MAREA, showing time-series for biophysical and socioeconomic indicators as well as tables containing the magnitude and significance of the $\beta_2$ and $\beta_1$ coefficients. A table with the governance indicators is also provided and discussed.

The scorecard (**Fig. 4**) shows that, overall, the reserve has a good performance achieving a general score of 63.6% of positive indicators. All category-level scores were also high, with values of 66.7%, 60%, and 75% positive indicators for Biophysical, Socioeconomic and Governance, respectively.

@@

Within the biophysical indicators, the greatest effect of the reserve was observed for snail and cucumber densities, with values of $\beta_3 = `r formatC(bio_results_summary$estimate[11], digits = 2, format = "f")`$ (*p* < 0.05) and $\beta_3 = `r formatC(bio_results_summary$estimate[12], digits = 2, format = "f")`$ (*p* < 0.05), respectively. Urchin and lobster densities also increased by $\beta_3 = `r formatC(bio_results_summary$estimate[10], digits = 2, format = "f")`$ and $\beta_3 = `r formatC(bio_results_summary$estimate[9], digits = 2, format = "f")`$ respectively. However, changes in urchin densities were only significant at *p* < 0.1, and not significant for lobster (*p* > 0.1). Shannon's diversity index ($\beta_3 = `r formatC(bio_results_summary$estimate[6], digits = 2, format = "f")`$) and richness $\beta_3 = `r formatC(bio_results_summary$estimate[7], digits = 2, format = "f")`$ showed a significant decrease (*p* < 0.05). Fish indicators showed no significant change (*p* > 0.1), with negative trends for Shannon's diversity index and fish species richness and positive trends for density, biomass, and mean trophic level. Changes through time for these indicators are presented in **Figure 5**, and a summary of $\beta_3$ coefficients is provided in **Table 3**.

In the case of socioeconomic indicators, total landings were on average 64.20 tones higher (*p* > 0.1) after the implementation of the reserves. Total income was \$10,344.85 (*p* < 0.05) thousands of Mexican Pesos (MXP) higher after the implementation of the reserves. On avergae, lobster and cucumber landings and increased, while urchin and snail landings and income decreased. **Figure 6** presents the changes in this indicators through time, and **Table 4** summarizes this information.

The results identified by our framework and MAREA are consistent to those previously published [*e.g.* @micheli_2012-EU; @rossetto_2015-V0]. Invertebrates show increased densities within the reserve, as compared to the control site, providing evidence that the reserve is effectively protecting these species.  While we are not able to infer causality from the socioeconomic analysis, we are able to identify important trends. For example, snail and urchin density have significantly increased within the reserve, but their landings and income have decreased. The opposite is observed for lobster and cucumber, which have shown increases in densities, landing, and income. While further information on market behaviour of each species is needed, these results provide insights into the state of the reserve, as well as the associated fisheries.

In terms of governance, it is evident that the community is strongly organized, which is a cause of their success. The first point of success is the existance of a fishing cooeprative that is also affiliated to federation. These polycentric governance structures allow different levels of organization that foster communication and cooperation; federations also provide bargain power with governments [@espinosaromero_2014-PY; @finkbeiner_2015-87]. Fishers also have good management instruments. Access to the fishing resources they exploit is managed through permits and fishing quotas. Along with a stable number of fishers participating in extractive activities, these limit the total fishing effort effort applied. Additionally, their TURF promotes a sense of stewardship of their resources and incentivizes correct resource management [@mccay_2017-1m; @finkbeiner_2015-87]. Together, these structures enabled a participative, bottom-up process during the reserve design phase; Opinions of all fishing members -and often non-fishers, but community members- were included. Participation of community members in reserve survailance and yearly monitoring indicate commitment and interest, and allow informal communication of results to un-involved community members.

While the reserve has not been legaly recognized (as of this moment), the community has started the process to have the reserve be recognized as a fishery replenishment zone (*Zona de Refugio Pesquer*) under the corresponding Mexican norms [@nom049sagpesc_2014-V6]. Furthermore, the reserve is partially isolated from poaching activity and fishers have internal regulations pertaining the reserves. The low level of illegal fishing by members of the community and outsiders both inside and outside the reserve represents another indication of effectiveness. A summary of governance indicators is provided in **Table 5**.

@@

```{r table of bio results, results = "asis"}
bio_results_summary %>% 
  mutate(estimate = paste0(formatC(estimate, digits = 2, format = "f"),
                           " (",
                           formatC(std.error, digits = 2, format = "f"),
                           ")",
                           significance)) %>% 
  select(Indicator, estimate, tscore = statistic) %>% 
  kable(booktabs = T,
        caption = "Summary of average treatment effect of the La Plana / Las Cuevas marine reserve on biophysical indicators. Asterisks indicate significance level, with (*) indicating *p* < 0.1 and (**) *p* < 0.05.",
        digits = 2,
        col.names = c("Indicator", "Estimate (SD)", "t-score")) %>% 
  kable_styling()
```

```{r table of soc results, results = "asis"}
socio_results_summary %>% 
  mutate(estimate = paste0(formatC(estimate, digits = 2, format = "f"),
                           " (",
                           formatC(std.error, digits = 2, format = "f"),
                           ")",
                           significance)) %>% 
  select(Indicator, estimate, tscore = statistic) %>% 
  kable(booktabs = T,
        caption = "Summary of average treatment effect of the La Plana / Las Cuevas marine reserve on socioeconomic indicators. Asterisks indicate significance level, with (*) indicating *p* < 0.1 and (**) *p* < 0.05.",
        digits = 2,
        col.names = c("Indicator", "Estimate (SD)", "t-score")) %>% 
  kable_styling()
```

@@

```{r plots, fig.height = 6, fig.width = 6, fig.cap = "Plots for values of each biophysical indicator (y-axis) through time (x-axis). Red and blue correspond to the reserve and control sites, respectively. Black lines indicate yearly mean values, and ribbons indicate +/- 1 standard error. Dots are horizontally jittered to aid visualization."}
shannon_fish_plot <- fish_results$plot[[1]] %>% 
  my_theme(y.lab = "H'")

richness_fish_plot <- fish_results$plot[[2]] %>% 
  my_theme(y.lab = "S (spp/transect)")

density_fish_plot <- fish_results$plot[[3]] %>% 
  my_theme(y.lab = "D (org/transect)")

trophic_fish_plot <- fish_results$plot[[4]] %>% 
  my_theme(y.lab = "TL")

biomass_fish_plot <- fish_results$plot[[5]] %>% 
  my_theme(y.lab = "B (Kg/transect)")

@@

shannon_invert_plot <- invert_results$plot[[1]] %>% 
  my_theme(y.lab = "H'")

richness_invert_plot <- invert_results$plot[[2]] %>% 
  my_theme(y.lab = "S (spp/transect)")

density_invert_plot <- invert_results$plot[[3]] %>% 
  my_theme(y.lab = "D (org/transect)")

lobster_density_invert_plot <- invert_results$plot[[4]] %>% 
  my_theme(y.lab = "lobsters (org/transect)")

urchin_density_invert_plot <- invert_results$plot[[5]] %>% 
  my_theme(x.lab = "year", y.lab = "urchin (org/transect)")

snail_density_invert_plot <- invert_results$plot[[6]] %>% 
  my_theme(x.lab = "year", y.lab = "snail (org/transect)")

cucumber_density_invert_plot <- invert_results$plot[[7]] %>% 
  my_theme(x.lab = "year", y.lab = "cucumber (org/transect)")

@@

```{r time series socio, fig.height = 6, fig.width = 6, fig.cap = "Plots for values of each socioeconomic indicator (y-axis) through time (x-axis). Red and blue correspond to before and after the implementation of the reserve, respectively."}
landings_plot <- socio$plot[[1]] %>% 
  my_theme(y.lab = "Landings (Kg)")

income_plot <- socio$plot[[2]] %>% 
  my_theme(y.lab = "Income (MXN)")

landings_lobster_plot <- socio$plot[[3]] %>% 
  my_theme(y.lab = "Lobster landings")

landings_urchin_plot <- socio$plot[[4]] %>% 
  my_theme(y.lab = "Urchin landings")

landings_snail_plot <- socio$plot[[5]] %>% 
  my_theme(y.lab = "Snail landings")

landings_cucumber_plot <- socio$plot[[6]] %>% 
  my_theme(y.lab = "Cucumber landings")

income_lobster_plot <- socio$plot[[7]] %>% 
  my_theme(y.lab = "Lobster income")

income_urchin_plot <- socio$plot[[8]] %>% 
  my_theme(y.lab = "Urchin income")

income_snail_plot <- socio$plot[[9]] %>% 
  my_theme(y.lab = "Snail income")

income_cucumber_plot <- socio$plot[[10]] %>% 
  my_theme(y.lab = "Cucumber income")

>>>>

# Results

The results shown here intend to highlight the relevance and utility of the framework and tool, which automate the analysis and make it replicable.  While we highlight some of the general observed trends, we focus on the utility of the tool rather than on the case study.

The scorecard (Fig. 4) shows that, overall, the reserve achieves a general score of 63.6% of positive indicators. All category-level scores were also high, with values of 66.7%, 60%, and 75% positive indicators for Biological, Socioeconomic and Governance, respectively.

@@

Among the biological indicators, the greatest effect of the reserve was observed for snail and cucumber densities, with values of $\beta_2 = `r formatC(bio_results_summary$estimate[11], digits = 2, format = "f")`$ (*p* < 0.05) and $\beta_2 = `r formatC(bio_results_summary$estimate[12], digits = 2, format = "f")`$ (*p* < 0.05), respectively. Fish indicators showed no significant change (*p* > 0.1), with negative trends for Shannon's diversity index and fish species richness and positive trends for density, biomass, and mean trophic level. Changes through time for these indicators are presented in Figure 5, and a summary of $\beta_2$ coefficients is provided in Table 3.

In the case of socioeconomic indicators, total landings were on average 64.20 tones higher (*p* > 0.1) after the implementation of the reserves. Total income was \$10,344.85 (*p* < 0.05) thousands of Mexican Pesos (MXP) higher after the implementation of the reserves. On average, lobster and cucumber landings increased, while urchin and snail landings and income decreased. Figure 6 presents the changes in these indicators through time, and Table 4 summarizes these results.

In terms of governance, it is evident that the community is strongly organized, which is likely a driver of their success. The first point of success is the existence of a fishing cooperative that is also affiliated to a regional fishing cooperatives federation. These polycentric governance structures allow various levels of organization that foster communication and cooperation; federations also provide bargain power with governments [@espinosaromero_2014-PY; @finkbeiner_2015-87]. Fishers also have good management instruments. Access to the fishing resources they exploit is managed through permits and fishing quotas. Along with a stable number of fishers participating in extractive activities, these limit the total fishing effortapplied. Additionally, their TURF promotes a sense of stewardship of their resources and incentivizes correct resource management [@mccay_2017-1m; @finkbeiner_2015-87]. Together, these structures enabled a participative, bottom--up process during the reserve design phase; Opinions of all fishing members ---and often non-fishers, but community members--- were included. Participation of community members in reserve surveillance and yearly monitoring indicate commitment and interest, and allow informal communication of results to un-involved community members.

Furthermore, the reserve is partially isolated from poaching activity and fishers have internal regulations pertaining the reserves. The low level of illegal fishing by members of the community and outsiders both inside and outside the reserve represents another indication of effectiveness. A summary of governance indicators is provided in Table 5.

@@

```{r table of bio results, results = "asis"}
bio_results_summary %>% 
  mutate(estimate = paste0(formatC(estimate, digits = 2, format = "f"),
                           " (",
                           formatC(std.error, digits = 2, format = "f"),
                           ")",
                           significance)) %>% 
  select(Indicator, estimate, tscore = statistic) %>% 
  kable(booktabs = T,
        caption = "Summary of average treatment effect of the reserve on biological indicators. Asterisks indicate significance level, with (*) indicating p < 0.1 and (**) p < 0.05.",
        digits = 2,
        col.names = c("Indicator", "Estimate (SD)", "t-score")) %>% 
  kable_styling()
```

```{r table of soc results, results = "asis"}
socio_results_summary %>% 
  mutate(estimate = paste0(formatC(estimate, digits = 2, format = "f"),
                           " (",
                           formatC(std.error, digits = 2, format = "f"),
                           ")",
                           significance)) %>% 
  select(Indicator, estimate, tscore = statistic) %>% 
  kable(booktabs = T,
        caption = "Summary of differences in socioeconomic indicators before and after the implementation of the reserve. Asterisks indicate significance level, with (*) indicating p < 0.1 and (**) p < 0.05.",
        digits = 2,
        col.names = c("Indicator", "Estimate (SD)", "t-score")) %>% 
  kable_styling()
```

@@

```{r plots, fig.height = 6, fig.width = 6, fig.cap = "Plots for values of each biological indicator (y-axis) through time (x-axis). Red and blue correspond to the reserve and control sites, respectively. Black lines indicate yearly mean values, and ribbons indicate +/- 1 standard error. Dots are horizontally jittered to aid visualization. This figure contains information for fish shannon diversity index (a), fish species richness (b), fish density (c), fish trophic level (d), fish biomass (e), invertebrate shannon diversity index (f), invertebrate species richness (g), invertebrate density (h), lobster density (i), urchin density (j), snail density (k), and cucumber density (l)."}

shannon_fish_plot <- fish_results$plot[[1]] %>% 
  my_theme(y.lab = "H'", text.x = 2005, text.y = 2.75, text.label = "a")

richness_fish_plot <- fish_results$plot[[2]] %>% 
  my_theme(y.lab = "S (spp./transect)", text.x = 2005, text.y = 9.5, text.label = "b")

density_fish_plot <- fish_results$plot[[3]] %>% 
  my_theme(y.lab = "D (org/transect)", text.x = 2005, text.y = 175, text.label = "c")

trophic_fish_plot <- fish_results$plot[[4]] %>% 
  my_theme(y.lab = "TL", text.x = 2005, text.y = 4.05, text.label = "d")

biomass_fish_plot <- fish_results$plot[[5]] %>% 
  my_theme(y.lab = "B (Kg/transect)", text.x = 2005, text.y = 200, text.label = "e")

@@

shannon_invert_plot <- invert_results$plot[[1]] %>% 
  my_theme(y.lab = "H'", text.x = 2005, text.y = 3.2, text.label = "f")

richness_invert_plot <- invert_results$plot[[2]] %>% 
  my_theme(y.lab = "S (spp./transect)", text.x = 2005, text.y = 14, text.label = "g")

density_invert_plot <- invert_results$plot[[3]] %>% 
  my_theme(y.lab = "D (org/transect)", text.x = 2005, text.y = 950, text.label = "h")

lobster_density_invert_plot <- invert_results$plot[[4]] %>% 
  my_theme(y.lab = "D (org/transect)", text.x = 2005, text.y = 275, text.label = "i")

urchin_density_invert_plot <- invert_results$plot[[5]] %>% 
  my_theme(x.lab = "", y.lab = "D (org/transect)", text.x = 2005, text.y = 50, text.label = "j")

snail_density_invert_plot <- invert_results$plot[[6]] %>% 
  my_theme(x.lab = "year", y.lab = "D (org/transect)", text.x = 2005, text.y = 700, text.label = "k")

cucumber_density_invert_plot <- invert_results$plot[[7]] %>% 
  my_theme(x.lab = "", y.lab = "D (org/transect)", text.x = 2005, text.y = 32, text.label = "l")

@@

```{r time series socio, fig.height = 6, fig.width = 6, fig.cap = "Plots for values of each socioeconomic indicator (y-axis) through time (x-axis). Red and blue correspond to before and after the implementation of the reserve, respectively. This figure contains information for total landings (a), total income (b), lobster landings (c), urchin landings (d), snail landings (e), cucumber landings (f), lobster income (g), urchin income (h), snail income (i), and cucumber income (j)."}

landings_plot <- socio$plot[[1]] %>% 
  my_theme(y.lab = "Landings (Kg)", text.x = 2001, text.y = 1150, text.label = "a", bio = F)

income_plot <- socio$plot[[2]] %>% 
  my_theme(y.lab = "Income (MXP)", text.x = 2001, text.y = 35000, text.label = "b", bio = F)

landings_lobster_plot <- socio$plot[[3]] %>% 
  my_theme(y.lab = "Landings (Kg)", text.x = 2001, text.y = 175, text.label = "c", bio = F)

landings_urchin_plot <- socio$plot[[4]] %>% 
  my_theme(y.lab = "Landings (Kg)", text.x = 2001, text.y = 55, text.label = "d", bio = F)

landings_snail_plot <- socio$plot[[5]] %>% 
  my_theme(y.lab = "Landings (Kg)", text.x = 2001, text.y = 300, text.label = "e", bio = F)

landings_cucumber_plot <- socio$plot[[6]] %>% 
  my_theme(y.lab = "Landings (Kg)", text.x = 2001, text.y = 58, text.label = "f", bio = F)

income_lobster_plot <- socio$plot[[7]] %>% 
  my_theme(y.lab = "Income (MXP)", text.x = 2001, text.y = 35000, text.label = "g", bio = F)

income_urchin_plot <- socio$plot[[8]] %>% 
  my_theme(y.lab = "Income (MXP)", text.x = 2001, text.y = 10000, text.label = "h", bio = F)

income_snail_plot <- socio$plot[[9]] %>% 
  my_theme(y.lab = "Income (MXP)", text.x = 2001, text.y = 2100, text.label = "i", bio = F)

income_cucumber_plot <- socio$plot[[10]] %>% 
  my_theme(y.lab = "Income (MXP)", text.x = 2001, text.y = 1500, text.label = "j", bio = F)
